# Analysis

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

* Answers to the research questions
* Different methods considered
* Competing approaches
* Justifications

### 1.Do European countries that have higher GDPs really have lower diabetes prevalence ?

This first question serves more as a control, since we learned during our research prior to our project that countries with higher GDPs tend to have lower diabetes rates. Indeed, we can observe that in the EDA. 

It is important to note that, when we try to fit a linear model on these variables and observe correlations over all observations, we see that these relationships are not significant at all.

```{r}
#with outliers
cor(GDP_diabetes_cal$avg_gdp, GDP_diabetes_cal$prop_women_diabetes)
cor(GDP_diabetes_cal$avg_gdp, GDP_diabetes_cal$prop_men_diabetes)

lm_gdp_diabetes_women <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$prop_women_diabetes)

lm_gdp_diabetes_men <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$prop_men_diabetes)

kable(prettify(summary(lm_gdp_diabetes_women)))
kable(prettify(summary(lm_gdp_diabetes_men)))
```

However, once we exclude outliers, we see that the relationship is way more significant !

```{r}
#without outliers
cor(GDP_diabetes_cal2$avg_gdp, GDP_diabetes_cal2$prop_women_diabetes)
cor(GDP_diabetes_cal2$avg_gdp, GDP_diabetes_cal2$prop_men_diabetes)

lm_gdp_diabetes_women2 <- lm(GDP_diabetes_cal2$avg_gdp ~ GDP_diabetes_cal2$prop_women_diabetes)

lm_gdp_diabetes_men2 <- lm(GDP_diabetes_cal2$avg_gdp ~ GDP_diabetes_cal2$prop_men_diabetes)

kable(prettify(summary(lm_gdp_diabetes_women2)))
kable(prettify(summary(lm_gdp_diabetes_men2)))
```

In the EDA section, we grouped countries in 3 categories, according to the relationship between the GDP and diabetes rate. Here, we confirmed statistically that a relationship exists between these 2 variables when we remove outliers. This therefore made us think that these countries could be categorized into clusters.

To determine the number of clusters we used the elbow method. This method examines the percentage of variance explained as a function of the number of clusters. It is based on the idea that a number of clusters should be chosen such that the addition of another cluster does not allow for a better modeling of the data. The percentage of variance explained by the clusters is plotted against the number of clusters.

```{r}
data_1 <- GDP_diabetes_cal[,-(1:2),drop=FALSE] #-> introduce the proportions later on ?
rownames(data_1) <- c("AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU","GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROU", "SVK", "SVN", "ESP", "SWE", "GBR", "CHE")

fviz_nbclust(data_1, kmeans, method = "wss") +
  geom_vline(xintercept = 3, linetype = 2) + # add line for better visualisation
  labs(subtitle = "Elbow method") # add subtitle

```

We therefore see from the graph above that the optimal number of clusters is 3. The allocation of countries according to their cluster is therefore as follows:

```{r}
set.seed(123)

model_cluster <- kmeans(data_1, centers = 3, nstart = 10)

Data_cluster <- data.frame(data_1,
  cluster = as.factor(model_cluster$cluster))

countries_cluster_1 <- filter(Data_cluster, cluster == "1")
countries_cluster_2 <- filter(Data_cluster, cluster == "2") 
countries_cluster_3 <- filter(Data_cluster, cluster == "3")

kable(countries_cluster_1)
kable(countries_cluster_2)
kable(countries_cluster_3)
```

Now let's plot these clusters to see the differences between them.

```{r}
res_kmeans <- cluster_analysis(data_1, n=3, method = "kmeans",iterations = 20)

predict(res_kmeans)

plot(res_kmeans)
```

We can therefore see that diabetes is indeed lower in clusters which contain the countries with higher GPDs. Let's see if this relationship is significant within each cluster :

```{r}
cor(countries_cluster_1$avg_gdp, countries_cluster_1$prop_women_diabetes)
cor(countries_cluster_1$avg_gdp, countries_cluster_1$prop_men_diabetes)
cor(countries_cluster_2$avg_gdp, countries_cluster_2$prop_women_diabetes)
cor(countries_cluster_2$avg_gdp, countries_cluster_2$prop_men_diabetes)
cor(countries_cluster_3$avg_gdp, countries_cluster_3$prop_women_diabetes)
cor(countries_cluster_3$avg_gdp, countries_cluster_3$prop_men_diabetes)

lm_gdp_diabetes_women_clust1 <- lm(countries_cluster_1$avg_gdp ~ countries_cluster_1$prop_women_diabetes)
lm_gdp_diabetes_men_clust1 <- lm(countries_cluster_1$avg_gdp ~ countries_cluster_1$prop_men_diabetes)
lm_gdp_diabetes_women_clust2 <- lm(countries_cluster_2$avg_gdp ~ countries_cluster_2$prop_women_diabetes)
lm_gdp_diabetes_men_clust2 <- lm(countries_cluster_2$avg_gdp ~ countries_cluster_2$prop_men_diabetes)
lm_gdp_diabetes_women_clust3 <- lm(countries_cluster_3$avg_gdp ~ countries_cluster_3$prop_women_diabetes)
lm_gdp_diabetes_men_clust3 <- lm(countries_cluster_3$avg_gdp ~ countries_cluster_3$prop_men_diabetes)

kable(prettify(summary(lm_gdp_diabetes_women_clust1)))
kable(prettify(summary(lm_gdp_diabetes_men_clust1)))
kable(prettify(summary(lm_gdp_diabetes_women_clust2)))
kable(prettify(summary(lm_gdp_diabetes_men_clust2)))
kable(prettify(summary(lm_gdp_diabetes_women_clust3)))
kable(prettify(summary(lm_gdp_diabetes_men_clust3)))
```

Clustering here doesn't necessarily help us answer the research question. We can confirm with the previous statistical test without outliers that overall, the GDP of a country and its diabetes rate are negatively correlated. 

However, the clusters defined above could help us answer our other research questions.

### 2.Do European countries that have higher GDPs consume less calories ?

As mentioned in the first point, countries with a higher GDP tend to have a lower diabetes rate, which could potentially be explained by the consumption of fewer calories.

But is there a real correlation between these two variables ? Let's check :

```{r}
cor(GDP_diabetes_cal$avg_gdp,GDP_diabetes_cal$total_consumption)

lm_gdp_total_cal <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$total_consumption)

kable(prettify(summary(lm_gdp_total_cal)))
```

Neither the correlation between these two variables nor the linear regression is significant. However, it would be interesting to look further. First, we can see if using the GDP per person instead of the total average makes a difference.

```{r}
ggplot() + 
  geom_point(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = total_consumption), color = "orange") +
  geom_smooth(mapping = aes(x = GDP_diabetes_cal$gdp_per_person, y=GDP_diabetes_cal$total_consumption), color = "orange", method = "lm") +
  labs(x = "GDP per capita", y= "Calories")
```

The plot looks more or less the same as the one in the EDA section without the outliers. But is the relationship with this new variable statistically significant ?

```{r}
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$total_consumption)

lm_gdp_pp_total_cal <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$total_consumption)

kable(prettify(summary(lm_gdp_pp_total_cal)))
```

We see that the correlation with the variable `gdp_per_person`is higher now and the significance of the parameter in the regression, even though not high enough, increased.

Next we can proceed with an analysis within clusters, with the ones defined in the first question.

```{r}
cor(countries_cluster_1$gdp_per_person, countries_cluster_1$total_consumption)
cor(countries_cluster_2$gdp_per_person, countries_cluster_2$total_consumption)
cor(countries_cluster_3$gdp_per_person, countries_cluster_3$total_consumption)

lm_gdp_total_cal_clust1 <- lm(countries_cluster_1$gdp_per_person ~ countries_cluster_1$total_consumption)
lm_gdp_total_cal_clust2 <- lm(countries_cluster_2$gdp_per_person ~ countries_cluster_2$total_consumption)
lm_gdp_total_cal_clust3 <- lm(countries_cluster_3$gdp_per_person ~ countries_cluster_3$total_consumption)

kable(prettify(summary(lm_gdp_total_cal_clust1)))
kable(prettify(summary(lm_gdp_total_cal_clust2)))
kable(prettify(summary(lm_gdp_total_cal_clust3)))
```
The results indicate that even within the clusters, these variables are not exactly correlated and the total calories consumed per person in a country is not a significant indicator of this country's wealth.

When we look at our cluster plot, in terms of total calorie intake it is also the 3rd cluster, which contains the richest countries considering GDP per person, that consumes the most calories. One might therefore think that calorie consumption is not the main reason why high-GDP countries have lower diabetes rates.

We are going to investigate this very issue of calorie consumption patterns that could be related to low diabetes rates. **-> à effacer ou question 4 ? **

```{r}
#trying to plot the consumption patterns
#library(plotly)
#library(dplyr)

#donut<- data.frame(GDP_diabetes_cal , cat=c("cal_prot_animal", "cal_carbs", "cal_prot_plant", "cal_fat"))

#GDP_diabetes_cal$total_consumption <- sapply(strsplit(rownames(GDP_diabetes_cal), " "), "[[", 1)

#donut <- donut %>% group_by(cat)
#donut <- donut %>% summarize(count = n())
#fig <- donut %>% plot_ly(labels = ~cat, values = ~count)
#fig <- fig %>% add_pie(hole = 0.6)
#fig <- fig %>% layout(title = "Donut charts to show consumption patterns",  showlegend = F,
                     # xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                     # yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

#fig

```



### 3.How do the macronutrients (animal protein/plant protein/fat/carbohydrates) consumed differ between richer and poorer countries ?

We observed during the EDA that richer countries seemed to consume more fat on average. Now we want to see if we can confirm this relationship, and observe if there isn't a correlation between the average GDP of a country and the consumption of other macronutrients too.

```{r}
cor(GDP_diabetes_cal$avg_gdp,GDP_diabetes_cal$cal_prot_animal)
cor(GDP_diabetes_cal$avg_gdp,GDP_diabetes_cal$cal_prot_plant)
cor(GDP_diabetes_cal$avg_gdp,GDP_diabetes_cal$cal_carbs)
cor(GDP_diabetes_cal$avg_gdp,GDP_diabetes_cal$cal_fat)


lm_gdp_cal_prot_animal <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$cal_prot_animal)
lm_gdp_cal_prot_plant <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$cal_prot_plant)
lm_gdp_cal_carbs <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$cal_carbs)
lm_gdp_cal_fat <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$cal_fat)

kable(prettify(summary(lm_gdp_cal_prot_animal)))
kable(prettify(summary(lm_gdp_cal_prot_plant)))
kable(prettify(summary(lm_gdp_cal_carbs)))
kable(prettify(summary(lm_gdp_cal_fat)))
```

The relationship between the average GDP of a country and the calories consumed from fat per person is the only macronutriment that is significant. For other macronutriments, there seems to be no correlation at all.

Let's see if considering the GDP per person instead of the total makes a difference.

```{r}
#with GDP per person
ggplot() + 
  geom_point(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = cal_fat), color = "blue") +
  geom_smooth(mapping = aes(x = GDP_diabetes_cal$gdp_per_person, y=GDP_diabetes_cal$cal_fat), color = "blue", method = "lm") +
  labs(x = "GDP per capita", y= "Calories from fat")

ggplot() +
  geom_point(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = cal_carbs), color = "purple") +
  geom_smooth(mapping = aes(x = GDP_diabetes_cal$gdp_per_person, y=GDP_diabetes_cal$cal_carbs), color = "purple", method = "lm") +
  labs(x = "GDP per capita", y= "Calories from carbs")

ggplot() +
  geom_point(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = cal_prot_animal), color = "red") +  
  geom_smooth(mapping = aes(x = GDP_diabetes_cal$gdp_per_person, y=GDP_diabetes_cal$cal_prot_animal), color = "red", method = "lm") +
  labs(x = "GDP per capita", y= "Calories from animal protein")

ggplot() +
  geom_point(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = cal_prot_plant), color = "green") +
  geom_smooth(mapping = aes(x = GDP_diabetes_cal$gdp_per_person, y=GDP_diabetes_cal$cal_prot_plant), color = "green", method = "lm") +
  labs(x = "GDP per capita", y= "Calories from plant protein")

#correlations and linear regressions
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$cal_prot_animal)
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$cal_prot_plant)
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$cal_carbs)
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$cal_fat)


lm_gdp_pp_cal_prot_animal <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$cal_prot_animal)
lm_gdp_pp_cal_prot_plant <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$cal_prot_plant)
lm_gdp_pp_cal_carbs <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$cal_carbs)
lm_gdp_pp_cal_fat <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$cal_fat)

kable(prettify(summary(lm_gdp_pp_cal_prot_animal)))
kable(prettify(summary(lm_gdp_pp_cal_prot_plant)))
kable(prettify(summary(lm_gdp_pp_cal_carbs)))
kable(prettify(summary(lm_gdp_pp_cal_fat)))
```

**insert analysis here**

Another way to answer this research question can be to see the relationship between the wealth of a country and the proportions of the total calories consumed dedicated to each macronutrient.

```{r}
GDP_diabetes_cal <- mutate(GDP_diabetes_cal, proportion_animal_prot = cal_prot_animal/total_consumption, proportion_plant_prot = cal_prot_plant/total_consumption, proportion_carbs = cal_carbs/total_consumption, proportion_fat = cal_fat/total_consumption)

cor(GDP_diabetes_cal$avg_gdp,GDP_diabetes_cal$proportion_animal_prot)
cor(GDP_diabetes_cal$avg_gdp,GDP_diabetes_cal$proportion_plant_prot)
cor(GDP_diabetes_cal$avg_gdp,GDP_diabetes_cal$proportion_carbs)
cor(GDP_diabetes_cal$avg_gdp,GDP_diabetes_cal$proportion_fat)

lm_gdp_prop_animal_prot <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$proportion_animal_prot)
lm_gdp_prop_plant_prot <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$proportion_plant_prot)
lm_gdp_prop_carbs <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$proportion_carbs)
lm_gdp_prop_fat <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$proportion_fat)

kable(prettify(summary(lm_gdp_prop_animal_prot)))
kable(prettify(summary(lm_gdp_prop_plant_prot)))
kable(prettify(summary(lm_gdp_prop_carbs)))
kable(prettify(summary(lm_gdp_prop_fat)))
```

With proportions, correlation is higher than with calorie count and linear regression parameters a bit more significant but the relationships are still not strong enough.

```{r}
#with GDP per person
ggplot() + 
  geom_point(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = proportion_fat), color = "blue") +
  geom_smooth(mapping = aes(x = GDP_diabetes_cal$gdp_per_person, y=GDP_diabetes_cal$proportion_fat), color = "blue", method = "lm") +
  labs(x = "GDP per capita", y= "Calories from fat")

ggplot() +
  geom_point(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = proportion_carbs), color = "purple") +
  geom_smooth(mapping = aes(x = GDP_diabetes_cal$gdp_per_person, y=GDP_diabetes_cal$proportion_carbs), color = "purple", method = "lm") +
  labs(x = "GDP per capita", y= "Calories from carbs")

ggplot() +
  geom_point(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = proportion_animal_prot), color = "red") +  
  geom_smooth(mapping = aes(x = GDP_diabetes_cal$gdp_per_person, y=GDP_diabetes_cal$proportion_animal_prot), color = "red", method = "lm") +
  labs(x = "GDP per capita", y= "Calories from animal protein")

ggplot() +
  geom_point(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = proportion_plant_prot), color = "green") +
  geom_smooth(mapping = aes(x = GDP_diabetes_cal$gdp_per_person, y=GDP_diabetes_cal$proportion_plant_prot), color = "green", method = "lm") +
  labs(x = "GDP per capita", y= "Calories from plant protein")

#correlations and linear regressions
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$proportion_animal_prot)
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$proportion_plant_prot)
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$proportion_carbs)
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$proportion_fat)

lm_gdp_pp_prop_animal_prot <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$proportion_animal_prot)
lm_gdp_pp_prop_plant_prot <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$proportion_plant_prot)
lm_gdp_pp_prop_carbs <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$proportion_carbs)
lm_gdp_pp_prop_fat <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$proportion_fat)

kable(prettify(summary(lm_gdp_pp_prop_animal_prot)))
kable(prettify(summary(lm_gdp_pp_prop_plant_prot)))
kable(prettify(summary(lm_gdp_pp_prop_carbs)))
kable(prettify(summary(lm_gdp_pp_prop_fat)))
```

**insert analysis here**

```{r}
#reintroduce clustering including proportions ?

#rest of code needs to be modified
cor(countries_cluster_1$avg_gdp, countries_cluster_1$prop_women_diabetes)
cor(countries_cluster_1$avg_gdp, countries_cluster_1$prop_men_diabetes)
cor(countries_cluster_2$avg_gdp, countries_cluster_2$prop_women_diabetes)
cor(countries_cluster_2$avg_gdp, countries_cluster_2$prop_men_diabetes)
cor(countries_cluster_3$avg_gdp, countries_cluster_3$prop_women_diabetes)
cor(countries_cluster_3$avg_gdp, countries_cluster_3$prop_men_diabetes)

lm_gdp_diabetes_women_clust1 <- lm(countries_cluster_1$avg_gdp ~ countries_cluster_1$prop_women_diabetes)
lm_gdp_diabetes_men_clust1 <- lm(countries_cluster_1$avg_gdp ~ countries_cluster_1$prop_men_diabetes)
lm_gdp_diabetes_women_clust2 <- lm(countries_cluster_2$avg_gdp ~ countries_cluster_2$prop_women_diabetes)
lm_gdp_diabetes_men_clust2 <- lm(countries_cluster_2$avg_gdp ~ countries_cluster_2$prop_men_diabetes)
lm_gdp_diabetes_women_clust3 <- lm(countries_cluster_3$avg_gdp ~ countries_cluster_3$prop_women_diabetes)
lm_gdp_diabetes_men_clust3 <- lm(countries_cluster_3$avg_gdp ~ countries_cluster_3$prop_men_diabetes)

kable(prettify(summary(lm_gdp_diabetes_women_clust1)))
kable(prettify(summary(lm_gdp_diabetes_men_clust1)))
kable(prettify(summary(lm_gdp_diabetes_women_clust2)))
kable(prettify(summary(lm_gdp_diabetes_men_clust2)))
kable(prettify(summary(lm_gdp_diabetes_women_clust3)))
kable(prettify(summary(lm_gdp_diabetes_men_clust3)))
```

**after within cluster analysis**
Oddly enough, it seems that a higher consumption of animal protein could be related to the rate of diabetes, which would be counter-intuitive to M. Adeva-Andany's (2019) article "Dietary habits contribute to define the risk of type 2 diabetes in humans".


### 4.How do these differences relate to the diabetes prevalence in these countries ? What is the typical diet that can be observed in richer states that relates to lower diabetes prevalence ?

What we observed in the EDA seemed to not make sense as we were expecting a positive relationship between the total calories/calories consumed from fat and diabetes prevalence in a country. Let's see now concretely if there is any correlation between the calories consumed from different macronutrients and the diabetes rate.

```{r}
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$cal_prot_animal)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$cal_prot_animal)
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$cal_prot_plant)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$cal_prot_plant)
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$cal_carbs)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$cal_carbs)
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$cal_fat)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$cal_fat)
```

The correlations are quite significant for fat calories, animal protein and also for plant protein. Nevertheless, carbs would not have a huge effect on the rate of diabetes. Furthermore, it is interesting to note that the correlation is always higher for women. Perhaps malnutrition raises the risk of diabetes in women more than in men? We will continue this analysis by running linear regressions. 

```{r}
lm_diabetes_men_cal_prot_animal <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$cal_prot_animal)
lm_diabetes_women_cal_prot_animal <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$cal_prot_animal)
lm_diabetes_men_cal_prot_plant <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$cal_prot_plant)
lm_diabetes_women_cal_prot_plant <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$cal_prot_plant)
lm_diabetes_men_cal_carbs <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$cal_carbs)
lm_diabetes_women_cal_carbs <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$cal_carbs)
lm_diabetes_men_cal_fat <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$cal_fat)
lm_diabetes_women_cal_fat <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$cal_fat)

kable(prettify(summary(lm_diabetes_men_cal_prot_animal)))
kable(prettify(summary(lm_diabetes_women_cal_prot_animal)))
kable(prettify(summary(lm_diabetes_men_cal_prot_plant)))
kable(prettify(summary(lm_diabetes_women_cal_prot_plant)))
kable(prettify(summary(lm_diabetes_men_cal_carbs)))
kable(prettify(summary(lm_diabetes_women_cal_carbs)))
kable(prettify(summary(lm_diabetes_men_cal_fat)))
kable(prettify(summary(lm_diabetes_women_cal_fat)))
```
Only animal protein and fat are significant for women which will further support the view that malnutrition has a greater impact on the rate of diabetes in women than in men. We will now take the proportions to get a better idea of the relationship of macronutrients to diabetes rates.
```{r}
#proportions
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$proportion_animal_prot)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$proportion_animal_prot)
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$proportion_plant_prot)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$proportion_plant_prot)
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$proportion_carbs)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$proportion_carbs)
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$proportion_fat)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$proportion_fat)

lm_diabetes_men_prop_prot_animal <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$proportion_animal_prot)
lm_diabetes_women_prop_prot_animal <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$proportion_animal_prot)
lm_diabetes_men_prop_prot_plant <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$proportion_plant_prot)
lm_diabetes_women_prop_prot_plant <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$proportion_plant_prot)
lm_diabetes_men_prop_carbs <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$proportion_carbs)
lm_diabetes_women_prop_carbs <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$proportion_carbs)
lm_diabetes_men_prop_fat <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$proportion_fat)
lm_diabetes_women_prop_fat <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$proportion_fat)

kable(prettify(summary(lm_diabetes_men_prop_prot_animal)))
kable(prettify(summary(lm_diabetes_women_prop_prot_animal)))
kable(prettify(summary(lm_diabetes_men_prop_prot_plant)))
kable(prettify(summary(lm_diabetes_women_prop_prot_plant)))
kable(prettify(summary(lm_diabetes_men_prop_carbs)))
kable(prettify(summary(lm_diabetes_women_prop_carbs)))
kable(prettify(summary(lm_diabetes_men_prop_fat)))
kable(prettify(summary(lm_diabetes_women_prop_fat)))
```
All macronutrients except protein are highly significant for women, which confirms our hypothesis that malnutrition has a greater impact on the rate of diabetes in women than in men. 

We are going to investigate this very issue of calorie consumption patterns that could be related to low diabetes rate for each cluster. To see the patterns of each cluster we will average each cluster. 
```{r}
#means for each cluster
countries_cluster_1 <- countries_cluster_1[,-10]
cl1 <- summarize_all(countries_cluster_1, mean)
cl1 <- rbind(countries_cluster_1, cl1)
row.names(cl1)[12] <- ("Means")
kable(cl1[12,])

countries_cluster_2 <- countries_cluster_2[,-10]
cl2 <- summarize_all(countries_cluster_2, mean)
cl2 <- rbind(countries_cluster_2, cl2)
row.names(cl2)[17] <- ("Means")
kable(cl2[17,])

countries_cluster_3 <- countries_cluster_3[,-10]
cl3 <- summarize_all(countries_cluster_3, mean)
cl3 <- rbind(countries_cluster_3, cl3)
row.names(cl3)[3] <- ("Means")
kable(cl3[3,])
```
As we see from the cluster averages, cluster 2 has the highest rate of diabetes followed by cluster 1 and finally the cluster 3 with the lowest rate of diabetes. 
```{r}
slices <- c(cl1[12,5],cl1[12,6],cl1[12,7], cl1[12,8] )
lbls <- c("Animal Protein", "Plant Protein", "Carbs", "Fat")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices, labels = lbls,col=rainbow(length(lbls)), main="Pie Chart of Consumption Pattern of Cluster 1")

slices <- c(cl2[17,5],cl2[17,6],cl2[17,7], cl2[17,8] )
lbls <- c("Animal Protein", "Plant Protein", "Carbs", "Fat")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices, labels = lbls,col=rainbow(length(lbls)), main="Pie Chart of Consumption Pattern of Cluster 2")

slices <- c(cl3[3,5],cl3[3,6],cl3[3,7], cl3[3,8] )
lbls <- c("Animal Protein", "Plant Protein", "Carbs", "Fat")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices, labels = lbls,col=rainbow(length(lbls)), main="Pie Chart of Consumption Pattern of Cluster 3")

```
There is not much difference between the different patterns for the different clusters. However, it should be noted that carbs represent 53% of the diet of cluster 2 and that this food greatly influences the rate of diabetes, particularly among women.  Indeed, there is a 1% increase in diabetes for every 0.168 carbs consumed on average per capita with a significant p-value at 1%. The proportion of carbs should therefore be reduced to less than half of the total proportion to try to have a negative impact on the diabetes rate.

Strangely enough, the proportion of fat is lower in cluster 1 (33%) which has the highest diabetes rate which is therefore counter intuitive. But this result is consistent with the negative correlations of fat with diabetes rate calculated above. It is therefore complicated to draw a conclusion regarding the proportion of fat to be consumed to reduce the diabetes rate. Our hypothesis is that even if cluster 1 consumes more fat, it is composed of rich countries that have more prevention measures and certainly better hospital infrastructure.
