# Analysis

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

* Answers to the research questions
* Different methods considered
* Competing approaches
* Justifications

## Our research questions

### 1.Do European countries that have higher GDPs really have lower diabetes prevalence ?

This first question serves more as a control, since we learned during our research prior to our project that countries with higher GDPs tend to have lower diabetes rates. Indeed, we can observe that in the EDA. 

But, when we try to fit a linear model on these variables and observe correlations we see that these relationships are not significant at all.

```{r}
#with outliers
cor(GDP_diabetes_cal$avg_gdp, GDP_diabetes_cal$prop_women_diabetes)
cor(GDP_diabetes_cal$avg_gdp, GDP_diabetes_cal$prop_men_diabetes)

lm_gdp_diabetes_women <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$prop_women_diabetes)

lm_gdp_diabetes_men <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$prop_men_diabetes)

summary(lm_gdp_diabetes_women)
summary(lm_gdp_diabetes_men)
```

However, once we exclude outliers, we see that the relationship is way more significant !

```{r}
#without outliers
cor(GDP_diabetes_cal2$avg_gdp, GDP_diabetes_cal2$prop_women_diabetes)
cor(GDP_diabetes_cal2$avg_gdp, GDP_diabetes_cal2$prop_men_diabetes)

lm_gdp_diabetes_women2 <- lm(GDP_diabetes_cal2$avg_gdp ~ GDP_diabetes_cal2$prop_women_diabetes)

lm_gdp_diabetes_men2 <- lm(GDP_diabetes_cal2$avg_gdp ~ GDP_diabetes_cal2$prop_men_diabetes)

summary(lm_gdp_diabetes_women2)
summary(lm_gdp_diabetes_men2)
```

#### LOOK SPECIFICALLY INTO RICH COUNTRIES

### 2.Do European countries that have higher GDPs consume less calories ?

As mentioned in the first point, countries with a higher GDP consume fewer calories which could be an explanation for the lower diabetes rate. Furthermore, we know from the EDA part that richer countries consume more fat on average but what is the real relationship between calorie consumption and GDP?

```{r}
#cor(GDP_diabetes_cal$avg_gdp, GDP_diabetes_cal$total_consumption)
#lm_cal =lm(GDP_diabetes_cal$country_code ~ GDP_diabetes_cal$cal_prot_animal + GDP_diabetes_cal$cal_prot_plant + #GDP_diabetes_cal$cal_carbs + GDP_diabetes_cal$cal_fat, data = GDP_diabetes_cal)

#summary(lm_cal)
```

Neither the correlation between these two variables nor the linear regression is significant. However, it will be interesting to look further.  To do this we can 
make clusters 

```{r}
rownames(GDP_diabetes_cal) <- c("AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU","GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROU", "SVK", "SVN", "ESP", "SWE", "GBR", "CHE")

data_1 <- GDP_diabetes_cal[,-(1:2),drop=FALSE] 
model <- kmeans(data_1, centers = 3)

# displays the class determined by
# the model for all observations:
model$cluster

Data_cluster <- data.frame(data_1,
  cluster = as.factor(model$cluster)
)

head(Data_cluster)

#prediction of cluster 
#install.packages('parameters')
#install.packages('see')
library(see)
library(parameters)

res_kmeans <- cluster_analysis(data_1,
                 n = 3,
                 method = "kmeans")

predict(res_kmeans) # get clusters

plot(res_kmeans)
```


```{r}
#trying to plot the consumption patterns
#library(plotly)
#library(dplyr)


#GDP_diabetes_cal$total_consumption <- sapply(strsplit(rownames(GDP_diabetes_cal), " "), "[[", 1)

#df <- GDP_diabetes_cal
#df <- df %>% group_by(total_consumption)
#df <- df %>% summarize(count = n())
#fig <- df %>% plot_ly(labels = ~total_consumption, values = ~count)
#fig <- fig %>% add_pie(hole = 0.6)
#fig <- fig %>% layout(title = "Donut charts using Plotly",  showlegend = F,
                      #xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      #yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

#fig


```



### 3.How do the macronutrients (animal protein/plant protein/fat/carbohydrates) consumed differ between richer and poorer governments ? And how do these differences relate to the diabetes prevalence in these countries ?

top 5, last 5 ?

parler des différents écarts entre macronutriments OU proportions par rapport aux total calories


### 4.What is the typical diet that can be observed in richer states that relates to lower diabetes prevalence ?

