# Analysis

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

* Answers to the research questions
* Different methods considered
* Competing approaches
* Justifications

## 1.Do European countries that have higher GDPs really have lower diabetes prevalence ?

This first question serves more as a control, since we learned during our research prior to our project that countries with higher GDPs tend to have lower diabetes rates. Indeed, we can observe that in the EDA. 

It is important to note that, when we try to fit a linear model on these variables and observe correlations over all observations, we see that these relationships are not significant at all.

```{r}
#with outliers

c0 <- matrix(c(cor(GDP_diabetes_cal$avg_gdp, GDP_diabetes_cal$prop_women_diabetes), cor(GDP_diabetes_cal$avg_gdp, GDP_diabetes_cal$prop_men_diabetes)))

rownames(c0) <- c("Between avg_gdp and women diabetes", "Between avg_gdp and men diabetes")

kbl(c0,
   caption = "**Correlation with outliers",
   align = c('cccccc')) %>%
   kable_classic( full_width = F, html_font = "Cambria") %>%
   kable_styling(position = "left", bootstrap_options = c('striped', 'hover', 'condensed')) 
```
<br><br>
```{r}

lm_gdp_diabetes_women <- lm(avg_gdp ~ prop_women_diabetes, data = GDP_diabetes_cal)

lm_gdp_diabetes_men <- lm(avg_gdp ~ prop_men_diabetes, data = GDP_diabetes_cal)

tab_model(
  lm_gdp_diabetes_women, show.ci = FALSE, show.stat = TRUE, show.se = TRUE,
  dv.labels = c("Diabetes vs Women diabetes"),
  CSS = list(
    css.width = '800px',
    css.centeralign = 'text-align: left;', 
    css.firsttablecol = 'font-weight: bold;',  
    css.summary = 'color: black;'
  )
)  
```
<br><br>
```{r}
tab_model(
  lm_gdp_diabetes_men, show.ci = FALSE, show.stat = TRUE, show.se = TRUE,
  dv.labels = c("Diabetes vs Men diabetes"),
  CSS = list(
    css.width = '500px',
    css.centeralign = 'text-align: left;', 
    css.firsttablecol = 'font-weight: bold;',  
    css.summary = 'color: black;'
  )
)


```

<br><br>

However, once we exclude outliers, we see that the relationship is way more significant !

```{r}
#without outliers

c_out <- matrix(c(cor(GDP_diabetes_cal2$avg_gdp, GDP_diabetes_cal2$prop_women_diabetes), cor(GDP_diabetes_cal2$avg_gdp, GDP_diabetes_cal2$prop_men_diabetes)))

rownames(c_out) <- c("Between avg_gdp and women diabetes", "Between avg_gdp and men diabetes")

kbl(c_out,
   caption = "**Correlation without outliers",
   align = c('cccccc')) %>%
   kable_classic(full_width = T, html_font = "Cambria") %>%
   kable_styling(bootstrap_options = c('striped', 'hover', 'condensed')) 
```
<br><br>
```{r}

lm_gdp_diabetes_women2 <- lm(avg_gdp ~ prop_women_diabetes, data = GDP_diabetes_cal2)

lm_gdp_diabetes_men2 <- lm(avg_gdp ~ prop_men_diabetes, data = GDP_diabetes_cal2)

tab_model(
  lm_gdp_diabetes_women2, show.ci = FALSE, show.stat = TRUE, show.se = TRUE,
  dv.labels = c("Diabetes vs Women diabetes </n> (without outliers"),
  CSS = list(
    css.width = '500px',
    css.centeralign = 'text-align: left;', 
    css.firsttablecol = 'font-weight: bold;',  
    css.summary = 'color: black;'
  )
)  
```
<br><br>
```{r}
tab_model(
  lm_gdp_diabetes_men2, show.ci = FALSE, show.stat = TRUE, show.se = TRUE,
  dv.labels = c("Diabetes vs Men diabetes </n> (without outliers"),
  CSS = list(
    css.width = '500px',
    css.centeralign = 'text-align: left;', 
    css.firsttablecol = 'font-weight: bold;',  
    css.summary = 'color: black;'
  )
)
```

In the EDA section, we grouped countries in 3 categories, according to the relationship between the GDP and diabetes rate. Here, we confirmed statistically that a relationship exists between these 2 variables when we remove outliers. This therefore made us think that these countries could be categorized into clusters.

To determine the number of clusters we used the elbow method. This method examines the percentage of variance explained as a function of the number of clusters. It is based on the idea that a number of clusters should be chosen such that the addition of another cluster does not allow for a better modeling of the data. The percentage of variance explained by the clusters is plotted against the number of clusters.

```{r, message=FALSE, warning=FALSE, include=FALSE}
data_1 <- GDP_diabetes_cal[,-(1:2),drop=FALSE] #-> introduce the proportions later on ?
rownames(data_1) <- c("AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU","GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROU", "SVK", "SVN", "ESP", "SWE", "GBR", "CHE")

fviz_nbclust(data_1, kmeans, method = "wss") +
  geom_vline(xintercept = 3, linetype = 2) + # add line for better visualisation
  labs(subtitle = "Elbow method")+ # add subtitle
  scale_y_continuous(labels = scales::label_number_si())
```

We therefore see from the graph above that the optimal number of clusters is 3. The allocation of countries according to their cluster is therefore as follows:

```{r, message=FALSE, warning=FALSE, include=FALSE}
set.seed(123)

model_cluster <- kmeans(data_1, centers = 3, nstart = 10)

Data_cluster <- data.frame(data_1, cluster = as.factor(model_cluster$cluster))

countries_cluster_1 <- filter(Data_cluster, cluster == "1")
countries_cluster_2 <- filter(Data_cluster, cluster == "2") 
countries_cluster_3 <- filter(Data_cluster, cluster == "3")

kbl(countries_cluster_1,
    caption = "**Table**: Cluster 1",
    align = c('cccccc')) %>%
   kable_classic(full_width = T, html_font = "Cambria") %>%
   kable_styling(bootstrap_options = c('striped', 'hover', 'condensed')) 
```
<br>
```{r}
kbl(countries_cluster_2,
    caption = "**Table**: Cluster 2",
    align = c('cccccc')) %>%
   kable_classic(full_width = T, html_font = "Cambria") %>%
   kable_styling(bootstrap_options = c('striped', 'hover', 'condensed'))
```
<br>
```{r}
kbl(countries_cluster_3,
    caption = "**Table**: Cluster 3",
    align = c('cccccc')) %>%
   kable_classic(full_width = T, html_font = "Cambria") %>%
   kable_styling(bootstrap_options = c('striped', 'hover', 'condensed'))
```
<br>

It is important to note that our 3rd cluster only contains 2 countries, and therefore when we analyse correlation and linear regression parameters, we will be mainly considering the overall nature of the relationships. (positive or negative link?)

Now let's plot these clusters to see the differences between them.

```{r, message=FALSE, warning=FALSE, include=FALSE}
res_kmeans <- cluster_analysis(data_1, n=3, method = "kmeans",iterations = 20)

plot(res_kmeans)


```

In the graph above,

* Group 1 = Cluster 3
* Group 2 = Cluster 2
* Group 3 = Cluster 1

We can therefore see that diabetes is indeed lower in clusters which contain the countries with higher GPDs. Let's see if this relationship is significant within each cluster :

```{r}
#cluster 1

c1 <- matrix(c(cor(countries_cluster_1$avg_gdp, countries_cluster_1$prop_women_diabetes), cor(countries_cluster_1$avg_gdp, countries_cluster_1$prop_men_diabetes)))

rownames(c1) <- c("Between avg_gdp and women diabetes", "Between avg_gdp and men diabetes")

kbl(c1,
   caption = "**Correlation for cluster 1",
   align = c('cccccc')) %>%
   kable_classic(full_width = T, html_font = "Cambria") %>%
   kable_styling(bootstrap_options = c('striped', 'hover', 'condensed')) 
```
<br><br>
```{r}
lm_gdp_diabetes_women_clust1 <- lm(avg_gdp ~ prop_women_diabetes, data = countries_cluster_1)
lm_gdp_diabetes_men_clust1 <- lm(avg_gdp ~ prop_men_diabetes, data = countries_cluster_1)

tab_model(
  lm_gdp_diabetes_women_clust1, show.ci = FALSE, show.stat = TRUE, show.se = TRUE,
  dv.labels = c("Diabetes vs Women diabetes (cluster 1)"),
  CSS = list(
    css.width = '500px',
    css.centeralign = 'text-align: left;', 
    css.firsttablecol = 'font-weight: bold;',  
    css.summary = 'color: black;'
  )
) 
```
<br><br>
```{r}
tab_model(
  lm_gdp_diabetes_men_clust1, show.ci = FALSE, show.stat = TRUE, show.se = TRUE,
  dv.labels = c("Diabetes vs Men diabetes (cluster 1)"),
  CSS = list(
    css.width = '500px',
    css.centeralign = 'text-align: left;', 
    css.firsttablecol = 'font-weight: bold;',  
    css.summary = 'color: black;'
  )
) 
#cluster 2

c2 <- matrix(c(cor(countries_cluster_2$avg_gdp, countries_cluster_2$prop_women_diabetes), cor(countries_cluster_2$avg_gdp, countries_cluster_2$prop_men_diabetes)))

rownames(c2) <- c("Between avg_gdp and women diabetes", "Between avg_gdp and men diabetes")

kbl(c2,
   caption = "**Correlation for cluster 2",
   align = c('cccccc')) %>%
   kable_classic(full_width = T, html_font = "Cambria") %>%
   kable_styling(bootstrap_options = c('striped', 'hover', 'condensed')) 
```
<br><br>
```{r}
lm_gdp_diabetes_women_clust2 <- lm(avg_gdp ~ prop_women_diabetes, data = countries_cluster_2)
lm_gdp_diabetes_men_clust2 <- lm(avg_gdp ~ prop_men_diabetes, data = countries_cluster_2)

tab_model(
  lm_gdp_diabetes_women_clust2, show.ci = FALSE, show.stat = TRUE, show.se = TRUE,
  dv.labels = c("Diabetes vs Women diabetes (cluster 2)"),
  CSS = list(
    css.width = '500px',
    css.centeralign = 'text-align: left;', 
    css.firsttablecol = 'font-weight: bold;',  
    css.summary = 'color: black;'
  )
) 
```
<br><br>

```{r, message=FALSE, warning=FALSE, include=FALSE}
tab_model(
  lm_gdp_diabetes_men_clust2, show.ci = FALSE, show.stat = TRUE, show.se = TRUE,
  dv.labels = c("Diabetes vs Men diabetes (cluster 2)"),
  CSS = list(
    css.width = '500px',
    css.centeralign = 'text-align: left;', 
    css.firsttablecol = 'font-weight: bold;',  
    css.summary = 'color: black;'
  )
) 

#cluster 3
c3 <- matrix(c(cor(countries_cluster_3$avg_gdp, countries_cluster_3$prop_women_diabetes), cor(countries_cluster_3$avg_gdp, countries_cluster_3$prop_men_diabetes)))

rownames(c3) <- c("Between avg_gdp and women diabetes", "Between avg_gdp and men diabetes")

kbl(c3,
   caption = "**Correlation for cluster 3",
   align = c('cccccc')) %>%
   kable_classic(full_width = T, html_font = "Cambria") %>%
   kable_styling(bootstrap_options = c('striped', 'hover', 'condensed')) 
```
<br><br>
```{r, message=FALSE, warning=FALSE, include=FALSE}
lm_gdp_diabetes_women_clust3 <- lm(avg_gdp ~ prop_women_diabetes, data = countries_cluster_3)
lm_gdp_diabetes_men_clust3 <- lm(avg_gdp ~ prop_men_diabetes, data = countries_cluster_3)

tab_model(
  lm_gdp_diabetes_women_clust3, show.ci = FALSE, show.stat = TRUE, show.se = TRUE,
  dv.labels = c("Diabetes vs Women diabetes (cluster 3)"),
  CSS = list(
    css.width = '500px',
    css.centeralign = 'text-align: left;', 
    css.firsttablecol = 'font-weight: bold;',  
    css.summary = 'color: black;'
  )
) 

```
<br><br>
```{r, message=FALSE, warning=FALSE, include=FALSE}
tab_model(
  lm_gdp_diabetes_men_clust3, show.ci = FALSE, show.stat = TRUE, show.se = TRUE,
  dv.labels = c("Diabetes vs Men diabetes (cluster 3)"),
  CSS = list(
    css.width = '500px',
    css.centeralign = 'text-align: left;', 
    css.firsttablecol = 'font-weight: bold;',  
    css.summary = 'color: black;'
  )
) 
```

To go further, we decided to represent on a cluster map to see if there was a difference between the north and the south for example.
```{r}

#EU_coord<- geojsonsf::geojson_sf('data/CNTR_RG_60M_2020_3035.geojson') # source of the geojson file #https://ec.europa.eu/eurostat/web/gisco/geodata/reference-data/administrative-units-statistical-units/countries

myCluster<-rbind.data.frame(countries_cluster_1, countries_cluster_2, countries_cluster_3)

EU_cluster <- c("AUT","BEL", "DNK", "FIN", "FRA", "DEU", "IRL", "ITA", "NLD", "SWE", "CHE", "BGR", "HRV", "CYP", "CZE", "EST", "GRC", "HUN", "LVA", "LTU", "MLT", "POL", "PRT", "ROU", "SVK", "SVN", "ESP", "LUX", "GBR")
CL_EU <- cbind(myCluster, EU_cluster)

get_eurostat_geospatial(resolution = 10, 
                        nuts_level = 0, 
                        year = 2016)

SHP_0 <- get_eurostat_geospatial(resolution = 10, 
                                 nuts_level = 0, 
                                 year = 2016)

eu <- c("SHQ", "AUT", "BEL", "NLD", "POL", "PRT", "DNK", "DEU", "ELL", "BGR", "CHE", "CYP", "ROU", "SRB","CZE", "EST", "HUN", "HRV", "SVN", "SVK", "SWE", "ITA", "TUR", "FIN", "NOR", "IRL", "ISL", "LIE", "LTU", "LUX", "LVA", "MNE", "MLT", "FRA")
SHP_0<- cbind(SHP_0, eu)
SHP_0 <- SHP_0[-c(1, 9, 15, 17, 24, 26, 28, 29, 33, 34), ]

#all_data_2 = merge(EU_coord, CL_EU, by.x = "ISO3_CODE", by.y = "EU_cluster")

nb_cluster <-as.numeric(levels(CL_EU$cluster))[CL_EU$cluster]
CL_EU <- cbind(CL_EU, nb_cluster)

leaflet(SHP_0) %>%
  setView(lng = 15, lat = 50, zoom = 4) %>% 
  addTiles() %>% 
  addPolygons(color = "black",
              weight = 1,
              fillOpacity = ifelse(  
                test = CL_EU$cluster == 1,  
                yes = 0.5,  
                no = 0  
              ),
              fillColor = "red")
```

The first cluster is therefore located more in the centre and north of Europe.
```{r}
leaflet(SHP_0) %>%
  setView(lng = 15, lat = 50, zoom = 4) %>% 
  addTiles() %>% 
  addPolygons(color = "black",
              weight = 1,
              fillOpacity = ifelse(  
                test = CL_EU$cluster == 2,  
                yes = 0.5,  
                no = 0  
              ),
              fillColor = "blue")
```
The second cluster would be located more to the southwest of Europe
```{r}
leaflet(SHP_0) %>%
  setView(lng = 15, lat = 50, zoom = 4) %>% 
  addTiles() %>% 
  addPolygons(color = "black",
              weight = 1,
              fillOpacity = ifelse(  
                test = CL_EU$cluster == 3,  
                yes = 0.5,  
                no = 0  
              ),
              fillColor = "orange")
```
And the third cluster represents only two countries so it is hard to give it a region. 

Clustering here doesn't necessarily help us answer the research question. We can confirm with the previous statistical test without outliers that overall, the GDP of a country and its diabetes rate are negatively correlated. 

However, the clusters defined above could help us answer our other research questions.


## 2.Do European countries that have higher GDPs consume less calories ?

As mentioned in the first point, countries with a higher GDP tend to have a lower diabetes rate, which could potentially be explained by the consumption of fewer calories.

But is there a real correlation between these two variables ? Let's check :

```{r}
c4 <- matrix(c(cor(GDP_diabetes_cal$avg_gdp,GDP_diabetes_cal$total_consumption)))

rownames(c4) <- c("Between avg_gdp and total calories")

kbl(c4,
   caption = "**Correlation",
   align = c('cccccc')) %>%
   kable_classic(full_width = T, html_font = "Cambria") %>%
   kable_styling(bootstrap_options = c('striped', 'hover', 'condensed')) 

lm_gdp_total_cal <- lm(avg_gdp ~ total_consumption, data = GDP_diabetes_cal)

tab_model(
  lm_gdp_total_cal, show.ci = FALSE, show.stat = TRUE, show.se = TRUE,
  dv.labels = c("avg_gdp vs total consumption"),
  CSS = list(
    css.width = '500px',
    css.centeralign = 'text-align: left;', 
    css.firsttablecol = 'font-weight: bold;',  
    css.summary = 'color: black;'
  )
)
```

Neither the correlation between these two variables nor the linear regression is significant. However, it would be interesting to look further. First, we can see if using the GDP per person instead of the total average makes a difference.

```{r}
ggplot() + 
  geom_point(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = total_consumption), color = "orange") +
  geom_smooth(mapping = aes(x = GDP_diabetes_cal$gdp_per_person, y=GDP_diabetes_cal$total_consumption), color = "orange", method = "lm") +
  labs(x = "GDP per capita", y= "Calories")
```

The plot looks more or less the same as the one in the EDA section without the outliers. But is the relationship with this new variable statistically significant ?

```{r}
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$total_consumption)

lm_gdp_pp_total_cal <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$total_consumption)

kable(prettify(summary(lm_gdp_pp_total_cal)))
```

We see that the correlation with the variable `gdp_per_person`is higher now and the significance of the parameter in the regression, even though not high enough, increased.

Next we can proceed with an analysis within clusters, with the ones defined in the first question.

```{r}
#cluster 1
cor(countries_cluster_1$gdp_per_person, countries_cluster_1$total_consumption)

lm_gdp_total_cal_clust1 <- lm(countries_cluster_1$gdp_per_person ~ countries_cluster_1$total_consumption)

kable(prettify(summary(lm_gdp_total_cal_clust1)))


#cluster 2
cor(countries_cluster_2$gdp_per_person, countries_cluster_2$total_consumption)

lm_gdp_total_cal_clust2 <- lm(countries_cluster_2$gdp_per_person ~ countries_cluster_2$total_consumption)

kable(prettify(summary(lm_gdp_total_cal_clust2)))


#cluster 3
cor(countries_cluster_3$gdp_per_person, countries_cluster_3$total_consumption)

lm_gdp_total_cal_clust3 <- lm(countries_cluster_3$gdp_per_person ~ countries_cluster_3$total_consumption)

kable(prettify(summary(lm_gdp_total_cal_clust3)))
```

The results indicate that even within the clusters, these variables are not exactly correlated and the total calories consumed per person in a country is not a significant indicator of this country's wealth.

When we look at our cluster plot, in terms of total calorie intake it is also the 1st cluster, which contains the richest countries considering GDP per capita, that consumes the most calories. One might therefore think that calorie consumption is not the main reason why high-GDP countries have lower diabetes rates.


## 3.How do the macronutrients (animal protein/plant protein/carbohydrates/fat) consumed differ between richer and poorer countries ?

We observed during the EDA that richer countries seemed to consume more fat on average. Now we want to see if we can confirm this relationship, and find out if there is a correlation between the average GDP of a country and the calories consumed for other macronutrients too.

Since the units of the variables "calories consumed from animal protein/plant protein/carbs/fat" are on a per capita basis, we think that including the GDP per capita in this part of the analysis could improve results. Additionally, we thought that another way of answering this research question could be to include the proportions of the calories consumed from different macronutrients on the total calories and see if it affects our analysis. 

```{r}
#adding proportions to the final table
GDP_diabetes_cal <- mutate(GDP_diabetes_cal, proportion_animal_prot = cal_prot_animal/total_consumption, proportion_plant_prot = cal_prot_plant/total_consumption, proportion_carbs = cal_carbs/total_consumption, proportion_fat = cal_fat/total_consumption)
```

With these new metrics introduced, we will plot these relationships again.

```{r}
#with GDP per person + calories
ggplot() + 
  geom_line(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = total_consumption), color = "orange") +
  geom_line(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = cal_fat), color = "blue") +
  geom_line(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = cal_carbs), color = "purple") +
  geom_line(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = cal_prot_animal), color = "red") +  
  geom_line(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = cal_prot_plant), color = "green") +
  labs(x = "GDP per capita", y= "Calories")

#with GDP per person + proportions
ggplot() + 
  geom_line(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = proportion_fat), color = "blue") +
  geom_line(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = proportion_carbs), color = "purple") +
  geom_line(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = proportion_animal_prot), color = "red") +  
  geom_line(data = GDP_diabetes_cal, aes(x = gdp_per_person, y = proportion_plant_prot), color = "green") +
  labs(x = "GDP per capita", y= "Proportion of calories")
```

It is also important to plot the clusters again to observe differences in the proportions of calories consumed.

```{r}
#clustering pt 2
data_2 <- mutate(data_1, GDP_diabetes_cal[,12:15])

set.seed(123)

model_cluster_bis <- kmeans(data_2, centers = 3, nstart = 10)

Data_cluster_bis <- data.frame(data_2, cluster = as.factor(model_cluster_bis$cluster))

countries_cluster_1bis <- filter(Data_cluster_bis, cluster == "1")
countries_cluster_2bis <- filter(Data_cluster_bis, cluster == "2") 
countries_cluster_3bis <- filter(Data_cluster_bis, cluster == "3")

res_kmeans_bis <- cluster_analysis(data_2, n=3, method = "kmeans",iterations = 20)

plot(res_kmeans_bis)
```

We will be referring to these graphs and this cluster plot in our analysis in the next sections.

### GDP - Animal protein

Let's start with by analysing the relationship between a country's wealth and it's consumption of animal protein. We saw at the EDA section that there didn't seem to be a trend between these 2 variables, overall the calories consumed from animal protein is very stable regardless of the average GDP of a country. We can confirm this by checking the correlation and by trying to fit a linear regression.

```{r}
#Average GDP - calories
cor(GDP_diabetes_cal$avg_gdp,GDP_diabetes_cal$cal_prot_animal)

lm_gdp_cal_prot_animal <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$cal_prot_animal)

kable(prettify(summary(lm_gdp_cal_prot_animal)))
```

Even though the correlation is positive, it is not very high. By looking at the summary of the linear regression, we can also see that the parameter is not significant. But what if we took the GDP per capita instead of the average GDP of a country ? The graph above seems to show again that the consumption of this macronutrient is very stable regardless of the GDP per capita of a country.

```{r}
#GDP per capita - calories
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$cal_prot_animal)

lm_gdp_pp_cal_prot_animal <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$cal_prot_animal)

kable(prettify(summary(lm_gdp_pp_cal_prot_animal)))
```

This time, the correlation is higher and the variable's estimate as a parameter is very significant ! Maybe the correlation can improve by considering proportion of calories instead of the calorie count.

```{r}
#GDP per capita - proportion
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$proportion_animal_prot)

lm_gdp_pp_prop_animal_prot <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$proportion_animal_prot)

kable(prettify(summary(lm_gdp_pp_prop_animal_prot)))
```

Here, the correlation is slightly lower than with the calorie count and the parameter's estimate is a bit less significant but still quite solid. Now let's see if we can find this positive relationship and better explanations within our defined clusters.

From our cluster plot above, we see that high GDP countries (clusters 1 and 3) tend to consume more calories of animal protein than low GDP countries, this is also the case for the proportions of calories consumed from animal protein. This follows what we have found so far. Are the correlations and linear regression in line with these observations ?

```{r}
#cluster 1
cor(countries_cluster_1bis$gdp_per_person, countries_cluster_1bis$proportion_animal_prot)

lm_gdp_pp_prop_animal_prot_clust1 <- lm(countries_cluster_1bis$gdp_per_person ~ countries_cluster_1bis$proportion_animal_prot)

kable(prettify(summary(lm_gdp_pp_prop_animal_prot_clust1)))


#cluster 2
cor(countries_cluster_2bis$gdp_per_person, countries_cluster_2bis$proportion_animal_prot)

lm_gdp_pp_prop_animal_prot_clust2 <- lm(countries_cluster_2bis$gdp_per_person ~ countries_cluster_2bis$proportion_animal_prot)

kable(prettify(summary(lm_gdp_pp_prop_animal_prot_clust2)))


#cluster 3
cor(countries_cluster_3bis$gdp_per_person, countries_cluster_3bis$proportion_animal_prot)

lm_gdp_pp_prop_animal_prot_clust3 <- lm(countries_cluster_3bis$gdp_per_person ~ countries_cluster_3bis$proportion_animal_prot)

kable(prettify(summary(lm_gdp_pp_prop_animal_prot_clust3)))
```

For each cluster, the correlation is positive and but still not so strong, and the estimate of the parameter for cluster 2 is significant.

In general, though it seems positive, we cannot conclude that there is a strong link between the wealth of a country and its animal protein consumption.

### GDP - Plant protein

Now let's move on to the analysis of the relationship between a country's wealth and it's consumption of plant protein. We saw at the EDA section, again, that there didn't seem to be a trend between these 2 variables. The calories consumed from plant protein is very stable as we can see from the horizontal line, regardless of the average GDP of a country. To confirm this, we will check the correlation.

```{r}
#Average GDP - calories
cor(GDP_diabetes_cal$avg_gdp,GDP_diabetes_cal$cal_prot_plant)

lm_gdp_cal_prot_plant <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$cal_prot_plant)

kable(prettify(summary(lm_gdp_cal_prot_plant)))
```

As we can see, the correlation is very low and when we try to fit a linear model over these 2 variables, the parameter estimation is not significant at all. Let's see if taking the GDP per capita makes a difference. In the graph above, the consumption of this macronutrient is very stable too.

```{r}
#GDP per capita - calories
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$cal_prot_plant)

lm_gdp_pp_cal_prot_plant <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$cal_prot_plant)

kable(prettify(summary(lm_gdp_pp_cal_prot_plant)))
```

The correlation improves when we consider the GDP per capita instead of the average. We also see that the relationship between the 2 variables is rather negative. However the parameter estimation is still not significant enough, even though it improved. Can we find a better explanation by using the proportion of calories rather than the calorie count of the plant protein consumed ?

```{r}
#GDP per capita - proportion
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$proportion_plant_prot)

lm_gdp_pp_prop_plant_prot <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$proportion_plant_prot)

kable(prettify(summary(lm_gdp_pp_prop_plant_prot)))
```

The correlation improved immensely ! Now we can confirm that the relationship is rather negative, even though the correlation is rather moderate than strong. But now our parameter estimation as significant as possible. How does this translate to our clusters ?

The cluster plot above indicates that in countries with higher GDPs like in cluster 1 and 3, the proportions of calories consumed from plant protein are lower than in countries with lower GDPs. Let's see if we can confirm these observations and see if these relationships are strong enough within each cluster.

```{r}
#cluster 1
cor(countries_cluster_1bis$gdp_per_person, countries_cluster_1bis$proportion_plant_prot)

lm_gdp_pp_prop_plant_prot_clust1 <- lm(countries_cluster_1bis$gdp_per_person ~ countries_cluster_1bis$proportion_plant_prot)

kable(prettify(summary(lm_gdp_pp_prop_plant_prot_clust1)))


#cluster 2
cor(countries_cluster_2bis$gdp_per_person, countries_cluster_2bis$proportion_plant_prot)

lm_gdp_pp_prop_plant_prot_clust2 <- lm(countries_cluster_2bis$gdp_per_person ~ countries_cluster_2bis$proportion_plant_prot)

kable(prettify(summary(lm_gdp_pp_prop_plant_prot_clust2)))


#cluster 3
cor(countries_cluster_3bis$gdp_per_person, countries_cluster_3bis$proportion_plant_prot)

lm_gdp_pp_prop_plant_prot_clust3 <- lm(countries_cluster_3bis$gdp_per_person ~ countries_cluster_3bis$proportion_plant_prot)

kable(prettify(summary(lm_gdp_pp_prop_plant_prot_clust3)))
```

We see that the correlations within clusters is weaker. And actually, even in countries with lower GDPs that belong to cluster 2, the relationship is rather negative. This indicates us that visual interpretation is not enough for analysis. We can maybe say that less wealthy countries that are more developed than those in the same cluster as them may consume more plant protein than more wealthy countries but they still consume less of this macronutrient than their cluster mates.

Therefore, we cannot conclude that there is a significant link between the wealth of a country and its plant protein consumption, although this relationship seems negative.

### GDP - Carbohydrates

Next, we will be evaluating the relationship of a country's GDP and its carbohydrate consumption. In the EDA graph, there isn't really trend in general between these 2 variables. The calories consumed from carbs is more or less stable for countries with an average GDP of around 500 billion dollars and more. Let's check the correlation.

```{r}
#Average GDP - calories
cor(GDP_diabetes_cal$avg_gdp,GDP_diabetes_cal$cal_carbs)

lm_gdp_cal_carbs <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$cal_carbs)

kable(prettify(summary(lm_gdp_cal_carbs)))
```

The correlation is negative and very low. This could maybe be explained by the fact that variance is very high at the left part of the graph where the average GDP is lower, up until 500 billion dollars. The parameter estimation is not significant either, this regression is not appropriate to explain this relationship.

But what if we compared the 2 metrics at the same unit level like in the previous 2 sections ? Our graph above indicates that there may be a slight decrease of the caloric intake of carbs per capita the more a country's GDP per capita is higher. Let's see if this is true.

```{r}
#GDP per capita - calories
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$cal_carbs)

lm_gdp_pp_cal_carbs <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$cal_carbs)

kable(prettify(summary(lm_gdp_pp_cal_carbs)))
```

Indeed, the correlation is higher now but it is still not strong enough. Our parameter is once again not significant to explain this new dependent variable either. But what if we changed our independent variable ? Because we see in the second graph above that this negative slope is even more accentuated when we take into account the proportion of the carbs consumed instead of their calorie count.

```{r}
#GDP per capita - proportion
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$proportion_carbs)

lm_gdp_pp_prop_carbs <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$proportion_carbs)

kable(prettify(summary(lm_gdp_pp_prop_carbs)))
```

The correlation between these 2 variables is way more important, we can even say that their correlation is moderately strong. And the parameter estimation when we fit a linear model is quite significant. Let's observe these relationships now within each cluster. According to our cluster plot, countries of cluster 1 and 3 (high GDP) consume a smaller proportion of carbs then those of cluster 2 (low GDP).

```{r}
#cluster 1
cor(countries_cluster_1bis$gdp_per_person, countries_cluster_1bis$proportion_carbs)

lm_gdp_pp_prop_carbs_clust1 <- lm(countries_cluster_1bis$gdp_per_person ~ countries_cluster_1bis$proportion_carbs)

kable(prettify(summary(lm_gdp_pp_prop_carbs_clust1)))


#cluster 2
cor(countries_cluster_2bis$gdp_per_person, countries_cluster_2bis$proportion_carbs)

lm_gdp_pp_prop_carbs_clust2 <- lm(countries_cluster_2bis$gdp_per_person ~ countries_cluster_2bis$proportion_carbs)

kable(prettify(summary(lm_gdp_pp_prop_carbs_clust2)))


#cluster 3
cor(countries_cluster_3bis$gdp_per_person, countries_cluster_3bis$proportion_carbs)

lm_gdp_pp_prop_carbs_clust3 <- lm(countries_cluster_3bis$gdp_per_person ~ countries_cluster_3bis$proportion_carbs)

kable(prettify(summary(lm_gdp_pp_prop_carbs_clust3)))
```

Surprisingly, with these correlations we see that the previous logic is inversed within clusters. For richer countries like those in cluster 1 and 3 the link between GDP and carb consumption is positive within their respective clusters, and for less rich countries like in those in cluster 2, it is negative. The estimate for the model parameter for cluster 2 is even very significant.

But overall, we cannot conclude that the link between the wealth of a country and its carb consumption is negative, as it is not strong enough.


### GDP - Fats

For the last part of our 3rd research question, let's analyse the relationship of a country's GDP and its fat consumption. Once again when we take a look at the graph at the EDA section, at the left side of the graph, there seems to be a slight increase. But the calories consumed from fat is more or less stable for countries with an average GDP of around 500 billion dollars and more, just like for the carbs. What does the correlation coefficient tell us ?


```{r}
#Average GDP - calories
cor(GDP_diabetes_cal$avg_gdp,GDP_diabetes_cal$cal_fat)

lm_gdp_cal_fat <- lm(GDP_diabetes_cal$avg_gdp ~ GDP_diabetes_cal$cal_fat)

kable(prettify(summary(lm_gdp_cal_fat)))
```

From the get-go, the correlation is moderately high and positive. The estimation of the model parameter is also quite significant. Does taking the GDP per capita change our results ? In the first graph above, the fat consumption seems to very slightly increase when a country's GDP is higher. Let's check.

```{r}
#GDP per capita - calories
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$cal_fat)

lm_gdp_pp_cal_fat <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$cal_fat)

kable(prettify(summary(lm_gdp_pp_cal_fat)))
```

Our hypothesis was right. Not only did the correlation between the 2 variables increase, but the significance of the parameter estimation also improved. And how does switching the calorie count with the proportion of fat consumed per capita affect the correlation with the GDP per capita ? On the second graph above, we see that, just like for carbs, the positive slope becomes a bit more accentuated.

```{r}
#GDP per capita - proportion
cor(GDP_diabetes_cal$gdp_per_person,GDP_diabetes_cal$proportion_fat)

lm_gdp_pp_prop_fat <- lm(GDP_diabetes_cal$gdp_per_person ~ GDP_diabetes_cal$proportion_fat)

kable(prettify(summary(lm_gdp_pp_prop_fat)))
```

This time, the correlation and the significance of the parameter goes down a little bit, even though not dramatically. Let's see once more what these relationships are like within each cluster. By looking at the cluster plot, we can hypothesize that wealthier countries of cluster 1 and 3 consume a bigger proportion of fat then those of cluster 2 that are less wealthy.

```{r}
#cluster 1
cor(countries_cluster_1bis$gdp_per_person, countries_cluster_1bis$proportion_fat)

lm_gdp_pp_prop_fat_clust1 <- lm(countries_cluster_1bis$gdp_per_person ~ countries_cluster_1bis$proportion_fat)

kable(prettify(summary(lm_gdp_pp_prop_fat_clust1)))


#cluster 2
cor(countries_cluster_2bis$gdp_per_person, countries_cluster_2bis$proportion_fat)

lm_gdp_pp_prop_fat_clust2 <- lm(countries_cluster_2bis$gdp_per_person ~ countries_cluster_2bis$proportion_fat)

kable(prettify(summary(lm_gdp_pp_prop_fat_clust2)))


#cluster 3
cor(countries_cluster_3bis$gdp_per_person, countries_cluster_3bis$proportion_fat)

lm_gdp_pp_prop_fat_clust3 <- lm(countries_cluster_3bis$gdp_per_person ~ countries_cluster_3bis$proportion_fat)

kable(prettify(summary(lm_gdp_pp_prop_fat_clust3)))
```

Just like for carbs, surprisingly, we see again that the previous logic is inversed within clusters by looking at these correlations. For high GDP countries like those in cluster 1 and 3 the link between GDP and fat consumption is negative within their respective clusters, and for low GDP countries like in those in cluster 2, it is positive. The estimate for the model parameter for cluster 2 is again quite significant. This suggests that among rich countries, the countries that are the most rich tend to consume less fat, and vice versa.

In general, even though it seems positive, we cannot confidently say that there is a definite link between the wealth of a country and its fat consumption.


## 4.How do these differences in macronutrients relate to the diabetes prevalence in these countries ? What is the typical diet that can be observed in richer states that relates to lower diabetes prevalence ?

What we observed in the EDA seemed to not make sense as we were expecting a positive relationship between the total calories/calories consumed from fat and diabetes prevalence in a country. Let's see now concretely if there is any correlation between the calories consumed from different macronutrients and the diabetes rate.

Oddly enough, it seems that a higher consumption of animal protein could be related to the rate of diabetes, which would be counter-intuitive to M. Adeva-Andany's (2019) article "Dietary habits contribute to define the risk of type 2 diabetes in humans". --> where to add ?


```{r}
#calories animal protein
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$cal_prot_animal)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$cal_prot_animal)

#calories plant protein
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$cal_prot_plant)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$cal_prot_plant)

#calories carbs
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$cal_carbs)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$cal_carbs)

#calories fat
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$cal_fat)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$cal_fat)
```

The correlations are quite significant for fat calories, animal protein and also for plant protein. Nevertheless, carbs would not have a huge effect on the rate of diabetes. Furthermore, it is interesting to note that the correlation is always higher for women. Perhaps malnutrition raises the risk of diabetes in women more than in men? We will continue this analysis by running linear regressions. 

```{r}
#calories animal protein
lm_diabetes_men_cal_prot_animal <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$cal_prot_animal)
lm_diabetes_women_cal_prot_animal <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$cal_prot_animal)

kable(prettify(summary(lm_diabetes_men_cal_prot_animal)))
kable(prettify(summary(lm_diabetes_women_cal_prot_animal)))


#calories plant protein
lm_diabetes_men_cal_prot_plant <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$cal_prot_plant)
lm_diabetes_women_cal_prot_plant <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$cal_prot_plant)

kable(prettify(summary(lm_diabetes_men_cal_prot_plant)))
kable(prettify(summary(lm_diabetes_women_cal_prot_plant)))


#calories carbs
lm_diabetes_men_cal_carbs <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$cal_carbs)
lm_diabetes_women_cal_carbs <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$cal_carbs)


kable(prettify(summary(lm_diabetes_men_cal_carbs)))
kable(prettify(summary(lm_diabetes_women_cal_carbs)))


#calories fat
lm_diabetes_men_cal_fat <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$cal_fat)
lm_diabetes_women_cal_fat <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$cal_fat)

kable(prettify(summary(lm_diabetes_men_cal_fat)))
kable(prettify(summary(lm_diabetes_women_cal_fat)))
```

Only animal protein and fat are significant for women which will further support the view that malnutrition has a greater impact on the rate of diabetes in women than in men. We will now take the proportions to get a better idea of the relationship of macronutrients to diabetes rates.

```{r}
#proportion animal protein
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$proportion_animal_prot)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$proportion_animal_prot)

#proportion plant protein
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$proportion_plant_prot)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$proportion_plant_prot)

#proportion carbs
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$proportion_carbs)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$proportion_carbs)

#proportion fat
cor(GDP_diabetes_cal$prop_men_diabetes,GDP_diabetes_cal$proportion_fat)
cor(GDP_diabetes_cal$prop_women_diabetes,GDP_diabetes_cal$proportion_fat)
```


```{r}
#proportion animal protein
lm_diabetes_men_prop_prot_animal <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$proportion_animal_prot)
lm_diabetes_women_prop_prot_animal <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$proportion_animal_prot)

kable(prettify(summary(lm_diabetes_men_prop_prot_animal)))
kable(prettify(summary(lm_diabetes_women_prop_prot_animal)))


#proportion plant protein
lm_diabetes_men_prop_prot_plant <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$proportion_plant_prot)
lm_diabetes_women_prop_prot_plant <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$proportion_plant_prot)

kable(prettify(summary(lm_diabetes_men_prop_prot_plant)))
kable(prettify(summary(lm_diabetes_women_prop_prot_plant)))


#proportion carbs
lm_diabetes_men_prop_carbs <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$proportion_carbs)
lm_diabetes_women_prop_carbs <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$proportion_carbs)

kable(prettify(summary(lm_diabetes_men_prop_carbs)))
kable(prettify(summary(lm_diabetes_women_prop_carbs)))


#proportion fat
lm_diabetes_men_prop_fat <- lm(GDP_diabetes_cal$prop_men_diabetes ~ GDP_diabetes_cal$proportion_fat)
lm_diabetes_women_prop_fat <- lm(GDP_diabetes_cal$prop_women_diabetes ~ GDP_diabetes_cal$proportion_fat)

kable(prettify(summary(lm_diabetes_men_prop_fat)))
kable(prettify(summary(lm_diabetes_women_prop_fat)))
```
All macronutrients except protein are highly significant for women, which confirms our hypothesis that malnutrition has a greater impact on the rate of diabetes in women than in men. 

We are going to investigate this very issue of calorie consumption patterns that could be related to low diabetes rate for each cluster. To see the patterns of each cluster we will average each cluster. 
```{r}
#means for each cluster
countries_cluster_1 <- countries_cluster_1[,-10]
cl1 <- summarize_all(countries_cluster_1, mean)
cl1 <- rbind(countries_cluster_1, cl1)
row.names(cl1)[12] <- ("Means")
kable(cl1[12,])

countries_cluster_2 <- countries_cluster_2[,-10]
cl2 <- summarize_all(countries_cluster_2, mean)
cl2 <- rbind(countries_cluster_2, cl2)
row.names(cl2)[17] <- ("Means")
kable(cl2[17,])

countries_cluster_3 <- countries_cluster_3[,-10]
cl3 <- summarize_all(countries_cluster_3, mean)
cl3 <- rbind(countries_cluster_3, cl3)
row.names(cl3)[3] <- ("Means")
kable(cl3[3,])
```
As we see from the cluster averages, cluster 2 has the highest rate of diabetes followed by cluster 1 and finally the cluster 3 with the lowest rate of diabetes. 
```{r}
slices <- c(cl1[12,5],cl1[12,6],cl1[12,7], cl1[12,8] )
lbls <- c("Animal Protein", "Plant Protein", "Carbs", "Fat")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices, labels = lbls,col=rainbow(length(lbls)), main="Pie Chart of Consumption Pattern of Cluster 1")

slices <- c(cl2[17,5],cl2[17,6],cl2[17,7], cl2[17,8] )
lbls <- c("Animal Protein", "Plant Protein", "Carbs", "Fat")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices, labels = lbls,col=rainbow(length(lbls)), main="Pie Chart of Consumption Pattern of Cluster 2")

slices <- c(cl3[3,5],cl3[3,6],cl3[3,7], cl3[3,8] )
lbls <- c("Animal Protein", "Plant Protein", "Carbs", "Fat")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices, labels = lbls,col=rainbow(length(lbls)), main="Pie Chart of Consumption Pattern of Cluster 3")

```

There is not much difference between the different patterns for the different clusters. However, it should be noted that carbs represent 53% of the diet of cluster 2 and that this food greatly influences the rate of diabetes, particularly among women.  Indeed, there is a 1% increase in diabetes for every 0.168 carbs consumed on average per capita with a significant p-value at 1%. The proportion of carbs should therefore be reduced to less than half of the total proportion to try to have a negative impact on the diabetes rate.

Strangely enough, the proportion of fat is lower in cluster 1 (33%) which has the highest diabetes rate which is therefore counter intuitive. But this result is consistent with the negative correlations of fat with diabetes rate calculated above. It is therefore complicated to draw a conclusion regarding the proportion of fat to be consumed to reduce the diabetes rate. Our hypothesis is that even if cluster 1 consumes more fat, it is composed of rich countries that have more prevention measures and certainly better hospital infrastructure.
