# Analysis

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/Code.R"))
```

* Answers to the research questions
* Different methods considered
* Competing approaches
* Justifications

## Our research questions
```
1.Do European countries that have higher GDPs really have lower diabetes prevalence ?

```{r echo= FALSE}
cor(GDP_diabetes_test$avg_gdp,GDP_diabetes_test$prop_men_diabetes)
cor(GDP_diabetes_test$avg_gdp,GDP_diabetes_test$prop_women_diabetes)

```
We see a very small correlation. It would be interesting to push the analysis a bit further to see the representation in a graph. 

```{r echo= FALSE}
plot(GDP_diabetes_test$avg_gdp,GDP_diabetes_test$prop_men_diabetes)
plot(GDP_diabetes_test$avg_gdp,GDP_diabetes_test$prop_women_diabetes)
```


2.Do European countries that have higher GDPs consume less calories ?
```{r echo= FALSE}
cor(GDP_diabetes_test$avg_gdp, GDP_diabetes_test$total_consumption)
plot(GDP_diabetes_test$avg_gdp, GDP_diabetes_test$total_consumption)
```


3.How do the proportions of macronutrients (animal protein/plant protein/fat/carbohydrates) consumed differ between richer and poorer governments ? And how do these differences relate to the diabetes prevalence in these countries ?
```{r}

```

4.What is the typical diet that can be observed in richer states that relates to lower diabetes prevalence ?
```{r}

```

