# Exploratory data analysis

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

* Mapping out the underlying structure
* Identifying the most important variables
* Univariate visualizations
* Multivariate visualizations
* Summary tables

First, even though we will be taking the means of the variables with which we are trying to answer our questions, it is interesting to observe their evolution in each country over time. We started with the GDP.

## GDP per country

```{r}
ggplot(data = GDP) +
  geom_point(mapping = aes(x = year, y = avg_gdp)) +
  facet_wrap(~ country_name, nrow = 4)
```


## Plotting GDP against Diabetes (Men & Women)
```{r}
#since there are diabetes for men and women i am not sure the coloring on this graph is entirely correct
ggplot(data = GDP_diabetes_cal) +
  geom_point(mapping = aes(x = avg_gdp, y = prop_men_diabetes, color = "blue")) +
  geom_point(mapping = aes(x = avg_gdp, y = prop_women_diabetes, color = "red"))
```

However, we see that apart of 5 outliers, our observations are mostly bunched up at the left of the graph. We decided to exclude these 5 observations, to see if we can observe a trend with the other countries. These outliers, as observed in the graph before, are the countries that had a big increase of GDP in the time period of 2000-2013.

```{r}
GDP_no_outlier <- arrange(GDP_summarized, desc(GDP_summarized$avg_gdp))
GDP_no_outlier <- GDP_no_outlier[-c(1:5),]

GDP_diabetes_cal2 <- left_join(GDP_no_outlier,Diabetes_EU_men, by="country_code")
GDP_diabetes_cal2 <- left_join(GDP_diabetes_cal2,Diabetes_EU_women, by="country_code")
GDP_diabetes_cal2 <- left_join(GDP_diabetes_cal2,Caloric_consumption, by="country_code")

#since there are diabetes for men and women i am not sure the coloring on this graph is entirely correct
ggplot(data = GDP_diabetes_cal2) +
  geom_point(mapping = aes(x = avg_gdp, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = avg_gdp, y = prop_men_diabetes, color = "blue"), method = "lm") +
  geom_point(mapping = aes(x = avg_gdp, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = avg_gdp, y = prop_women_diabetes, color = "red"), method = "lm")
```

Without the outliers, we can see a bit more clearly. Indeed, it seems that the richer a country is, the lesser it has a high diabetes rate among its population. But what about the 5 most rich countries in Europe ?



## Evolution of calories in each countries

First, we tried to see if there was a trend in the consumption of macro-nutrients in the 2000s for each country in our sample by plotting those evolution over time:
```{r}
#Plotting the evolution of the micronutrients consumption over the years for each country

cal_unique <- unique(daily_caloric, by =c('Entity')) # Sorting table by entity (unique)

for (i in 1:nrow(cal_unique)) { # Loop for each entity 
tmp <- cal_unique$Entity[i]     # Getting the entity name i

# Taking subset of data table for the chosen entity i 
tmp <- subset(daily_caloric,Entity == tmp, select = c("Year","Calories from animal protein (FAO (2017))",
                                                            "Calories from plant protein (FAO (2017))",
                                                            "Calories from fat (FAO (2017))",
                                                            "Calories from carbohydrates (FAO (2017))"))

# Melting the data by years (i.e. getting each value for the different rows for each years) 
tmp <- melt(tmp ,  id.vars = 'Year', variable.name = 'Series')

# Print plot
print(ggplot(tmp, aes(Year,value)) + geom_line(aes(colour = Series),size = 2) +
  labs(title = paste("Caloric consumption per year",cal_unique$Entity[i]),
       y = "Value", x = "Year"))
  
}
```

There do not seem to be any trends in the graphs above and diets are rather stable in each country, which is why we will take the average consumption for each macro-nutrient. We can however note that the 5 outliers mentioned before have a higher fat consumption than the countries with a smaller GDP. 

## GDP againt total calories

We now plot the different consumption and the total consumption of each macro-nutrients to see if there's a trend:

```{r}
ggplot() + 
  geom_line(data = GDP_diabetes_cal, aes(x = avg_gdp, y = total_consumption), color = "blue") +
  geom_line(data = GDP_diabetes_cal, aes(x = avg_gdp, y = cal_fat), color = "red") +
  geom_line(data = GDP_diabetes_cal, aes(x = avg_gdp, y = cal_carbs), color = "green") +
  geom_line(data = GDP_diabetes_cal, aes(x = avg_gdp, y = cal_prot_animal), color = "purple") +  
  geom_line(data = GDP_diabetes_cal, aes(x = avg_gdp, y = cal_prot_plant), color = "orange") +
  xlab('Caloric consumption by GDP') +
  ylab('Value')

```

We see that total consumption does not really change. We then plot the total consumption with each country's GDP.

```{r}
ggplot(GDP_diabetes_cal, aes(x=GDP_diabetes_cal$avg_gdp, y=GDP_diabetes_cal$total_consumption)) +
  geom_point() + # Show dots
  geom_text(
   label= (GDP_diabetes_cal$country_code),  
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T)+
 geom_smooth(mapping = aes(x = GDP_diabetes_cal$avg_gdp, y=GDP_diabetes_cal$total_consumption, color = "red"), method = "lm")
```
We always end up with these 5 outliers that have a higher than average GDP so if we remove them from the average we have:

```{r}
ggplot(GDP_diabetes_cal2, aes(x=GDP_diabetes_cal2$avg_gdp, y=GDP_diabetes_cal2$total_consumption)) +
  geom_point() + # Show dots
  geom_text(
    label= (GDP_diabetes_cal2$country_code),  
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T)+
geom_smooth(mapping = aes(x = GDP_diabetes_cal2$avg_gdp, y=GDP_diabetes_cal2$total_consumption, color = "red"), method = "lm")
```
Now we can see clearly that there's a trend.





## Evolution of Diabetes over time

To begin with, we tried to see if there was a trend in the evolution of diabetes prevalence in the 2000s for each country in our sample by plotting those evolution over time:

```{r}
#plotting the evolution of the prevalence of diabetes for each country 
Diabetes_EU2 <- Diabetes_EU[-c(2)] #remove ISO country code
Diabetes_unique <- unique(Diabetes_EU2[c('country')]) # Sorting table by country (unique)

for (i in 1:nrow(Diabetes_unique)) { 
tmp <- Diabetes_unique$country[i]     

# Taking subset of data table for the chosen country i 
tmp2 <- subset(Diabetes_EU2,country == tmp, select = c("sex","year","prop_diabetes"))

# Print plot for each country
print(ggplot(data = tmp2, mapping = aes(x = year, y =prop_diabetes, color =sex)) + geom_line()
  + labs(title = paste("Evolution of diabetes prevalence in",tmp), y="Diabetes prevalence") + scale_y_continuous(labels = scales::percent))
}
```

We see right away that the prevalence of diabetes is higher for man than women across all countries (there are however two exceptions : in Romania between 2000 and 2003 and Slovenia between 2000 and 2006). 
Moreover, we observe that we have a decrease over time for these countries :
* Belgium
* Denmark
* Finland
We have countries with a decrease over time for women and not for men:
* Austria
* Malta
* Netherlands
* Germany
* Italy
* Spain
* Switzerland
In other european countries, the prevalence of diabetes is increasing (at different pace) over time.

## Plotting Diabetes against each type of calories consumption

Now we want to see the diabetes prevalence against total calories consumption and also against each type calories consumed:

```{r}
#plotting diabetes against total calories consumption and each type of calories consumed
#diabetes vs total consumption
ggplot(data = GDP_diabetes_cal) +
  geom_point(mapping = aes(x = total_consumption, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = total_consumption, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = total_consumption, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = total_consumption, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs total calories per day", x="Total calories per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)
#diabetes vs calories from animal protein
ggplot(data = GDP_diabetes_cal) +
  geom_point(mapping = aes(x = cal_prot_animal, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = cal_prot_animal, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = cal_prot_animal, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = cal_prot_animal, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs daily calories from animal protein", x="Total calories from animal protein per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)
#diabetes vs calories from plant protein
ggplot(data = GDP_diabetes_cal) +
  geom_point(mapping = aes(x = cal_prot_plant, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = cal_prot_plant, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = cal_prot_plant, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = cal_prot_plant, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs daily calories from plant protein", x="Total calories from plant protein per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)

#diabetes vs calories from carbohydrates
ggplot(data = GDP_diabetes_cal) +
  geom_point(mapping = aes(x = cal_carbs, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = cal_carbs, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = cal_carbs, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = cal_carbs, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs daily calories from carbohydrates", x="Total calories from carbohydrates per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)
```
We can see a negative trend for the total consumption and the calories from animal protein. We can observe a positive trend against calories from plant protein. For protein from carbohydrates, we can see a slighty positive trend for women.

Now we want to see if we have different trend when we remove our 5 outliers: 

```{r}
ggplot(data = GDP_diabetes_cal2) +
  geom_point(mapping = aes(x = total_consumption, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = total_consumption, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = total_consumption, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = total_consumption, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs total calories per day (without outliers)", x="Total calories per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)
#diabetes vs calories from animal protein
ggplot(data = GDP_diabetes_cal2) +
  geom_point(mapping = aes(x = cal_prot_animal, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = cal_prot_animal, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = cal_prot_animal, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = cal_prot_animal, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs daily calories from animal protein (without outliers)", x="Total calories from animal protein per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)
#diabetes vs calories from plant protein
ggplot(data = GDP_diabetes_cal2) +
  geom_point(mapping = aes(x = cal_prot_plant, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = cal_prot_plant, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = cal_prot_plant, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = cal_prot_plant, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs daily calories from plant protein (without outliers)", x="Total calories from plant protein per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)

#diabetes vs calories from carbohydrates
ggplot(data = GDP_diabetes_cal2) +
  geom_point(mapping = aes(x = cal_carbs, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = cal_carbs, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = cal_carbs, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = cal_carbs, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs daily calories from carbohydrates (without outliers)", x="Total calories from carbohydrates per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)

```

Without our 5 outliers, we observe not much change in the trend of each type of calories consumed apart for carbohydrates where the trend changes for men and become slighly positive.
