# Exploratory data analysis

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

First, even though we will be taking the means of the variables with which we are trying to answer our questions, it is interesting to observe their evolution in each country over time. We started with the GDP.


## Evolution of GDP per country

```{r}
ggplot(data = GDP) +
  geom_point(mapping = aes(x = year, y = avg_gdp)) +
  facet_wrap(~ country_name, nrow = 4) +
  labs(x="Year",y="GDP")
```

We can see that the GDP of France, Germany, Italy, Spain and the United Kingdom had a significant increase between 2000 and 2008.


## Plotting GDP against Diabetes (Men & Women)

Now let's see if there is a relation between the GDP of a country and its diabetes prevalence. (men = blue, women = red)

```{r}
ggplot(data = GDP_diabetes_cal) +
  geom_point(mapping = aes(x = avg_gdp, y = prop_men_diabetes),color = "blue") +
  geom_point(mapping = aes(x = avg_gdp, y = prop_women_diabetes), color = "red") +
  labs(x="GDP",y="Diabetes rate")
```

We observe that apart of 5 outliers, our observations are mostly bunched up at the left of the graph. We decided to exclude these 5 observations, to see if we can observe a trend with the other countries. These outliers, as we can see on the graph before, are the countries that had a big increase of GDP in the time period of 2000-2013.

```{r}
GDP_no_outlier <- arrange(GDP_summarized, desc(GDP_summarized$avg_gdp))
GDP_no_outlier <- GDP_no_outlier[-c(1:5),]

GDP_diabetes_cal2 <- left_join(GDP_no_outlier,Diabetes_EU_men, by="country_code")
GDP_diabetes_cal2 <- left_join(GDP_diabetes_cal2,Diabetes_EU_women, by="country_code")
GDP_diabetes_cal2 <- left_join(GDP_diabetes_cal2,Caloric_consumption, by="country_code")

ggplot(data = GDP_diabetes_cal2) +
  geom_point(mapping = aes(x = avg_gdp, y = prop_men_diabetes), color = "blue") +
  geom_smooth(mapping = aes(x = avg_gdp, y = prop_men_diabetes), color = "blue", method = "lm") +
  geom_point(mapping = aes(x = avg_gdp, y = prop_women_diabetes), color = "red") +
  geom_smooth(mapping = aes(x = avg_gdp, y = prop_women_diabetes), color = "red", method = "lm") +
  labs(x="GDP",y="Diabetes rate")
```

Without the outliers, we can see a bit more clearly. Indeed, it seems that the richer a country is, the lesser it has a high diabetes rate among its population.


## Evolution of the consumption of macronutrients in calories per country

For the second table, we tried to see again if there was a trend in the consumption of different macro-nutrients in the 2000s for each country in our sample.

```{r}

# Plotting the evolution of the macronutrients consumption over the years for each country
cal_unique <- unique(daily_caloric, by =c('Entity')) # Sorting table by entity (unique)

for (i in 1:nrow(cal_unique)) { # Loop for each entity 
tmp <- cal_unique$Entity[i]     # Getting the entity name i

# Taking subset of data table for the chosen entity i 
tmp <- subset(daily_caloric,Entity == tmp, select = c("Year","Calories from animal protein (FAO (2017))","Calories from plant protein (FAO (2017))","Calories from fat (FAO (2017))","Calories from carbohydrates (FAO (2017))"))

# Melting the data by years (i.e. getting each value for the different rows for each year) 
tmp <- melt(tmp ,  id.vars = 'Year', variable.name = 'Series')

# Print plot

print(ggplot(tmp, aes(Year,value)) + geom_line(aes(colour = Series),size = 2) +
  labs(title = paste("Caloric consumption per year",cal_unique$Entity[i]),
       y = "Calories", x = "Year"))
  }



library(shiny)
ui <- tagList(
    tags$head(
        tags$title("HTML Button Test")
    ),
    tags$main(
        tags$h1("HTML Button Test"),
        tags$p("Click the button."),
        tags$button(
            id = "btn",
            class = "shiny-bound-input action-button",
            "My Button"
        ) 
    )
)


#one way 
#install.packages("plotly")

#library(ggplot2)
#library(plotly)
#p <- ggplot(tmp, aes(Year,value)) + geom_line(aes(colour = Series),size = 2) +
  #labs(title = paste("Caloric consumption per year",cal_unique$Entity[i]),
       #y = "Calories", x = "Year")
#ggplotly(p)

```

There do not seem to be any trends in the graphs above and diets seem rather stable in each country, which is why we will take the average consumption for each macro-nutrient. We can however note that the 5 outliers mentioned before tend to have a higher fat consumption than the countries with a smaller GDP. 


## Plotting GDP against total calories consumed

We then wanted to analyse the relation between a country's GDP and its individual consumption of each macronutrient as well as its total calorie consumption to see if there's a trend.(total calories = orange, fat = blue, carbohydrates = purple, animal protein = red, plant protein = green)

```{r}
ggplot() + 
  geom_line(data = GDP_diabetes_cal, aes(x = avg_gdp, y = total_consumption), color = "orange") +
  geom_line(data = GDP_diabetes_cal, aes(x = avg_gdp, y = cal_fat), color = "blue") +
  geom_line(data = GDP_diabetes_cal, aes(x = avg_gdp, y = cal_carbs), color = "purple") +
  geom_line(data = GDP_diabetes_cal, aes(x = avg_gdp, y = cal_prot_animal), color = "red") +  
  geom_line(data = GDP_diabetes_cal, aes(x = avg_gdp, y = cal_prot_plant), color = "green") +
  labs(x = "GDP", y= "Calories")
```

We see that the total calorie consumption does not really change. We wanted a close up on the relation between the total calorie consumption with the GDP for each country to see if we can spot outliers again, so we created other plots.

```{r}
ggplot(GDP_diabetes_cal, aes(x=GDP_diabetes_cal$avg_gdp, y=GDP_diabetes_cal$total_consumption)) +
  geom_point() + # Show dots
  geom_text(
   label= (GDP_diabetes_cal$country_code),  
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T)+
 geom_smooth(mapping = aes(x = GDP_diabetes_cal$avg_gdp, y=GDP_diabetes_cal$total_consumption), color = "red", method = "lm")+
   labs( x = "Average GDP", y = "Calories")
```

We end up again with these 5 outliers that have a higher than average GDP so if we remove them, we obtain the following plots :

```{r}
ggplot(GDP_diabetes_cal2, aes(x=GDP_diabetes_cal2$avg_gdp, y=GDP_diabetes_cal2$total_consumption)) +
  geom_point() +
  geom_text(
    label= (GDP_diabetes_cal2$country_code),  
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T)+
geom_smooth(mapping = aes(x = GDP_diabetes_cal2$avg_gdp, y=GDP_diabetes_cal2$total_consumption), color = "red", method = "lm")+
   labs(x = "Average GDP", y = "Calories")
```

Now we can more easily state that there's a trend. It appears that the higher a country's GDP, the higher the total calories consumed, contrary to our hypothesis.


## Evolution of Diabetes per country

Once again, we tried to see if the diabetes prevalence in each country changed over the years 2000-2013.

```{r}
# Plotting the evolution of the prevalence of diabetes for each country 
#Diabetes_EU2 <- Diabetes_EU[-c(2)] #remove ISO country code
#Diabetes_unique <- unique(Diabetes_EU2[c('country')]) # Sorting table by country (unique)

#for (i in 1:nrow(Diabetes_unique)) { 
#tmp <- Diabetes_unique$country[i]     

# Taking subset of data table for the chosen country i 
#tmp2 <- subset(Diabetes_EU2,country == tmp, select = c("sex","year","prop_diabetes"))

# Print plot for each country
#print(ggplot(data = tmp2, mapping = aes(x = year, y =prop_diabetes, color =sex)) + geom_line()
#  + labs(title = paste("Evolution of diabetes prevalence in",tmp), x= "Year", y="Diabetes prevalence") + scale_y_continuous(labels = scales::percent))
#}
```


We saw right away that the prevalence of diabetes is higher for man than women across all countries (there are however two exceptions : in Romania between 2000 and 2003 and Slovenia between 2000 and 2006).

We observed three different scenarios for countries that we selected:
A decrease over time of diabetes over time for:

* Belgium  
* Denmark  
* Finland  

We take Belgium as an example :
```{r}
# Plotting the evolution of the prevalence of diabetes for each country 
Diabetes_EU2 <- Diabetes_EU[-c(2)] #remove ISO country code
Diabetes_unique <- unique(Diabetes_EU2[c('country')]) # Sorting table by country (unique)


# Taking subset of data table for the chosen country
BelgiumPlot <- subset(Diabetes_EU2,country == 'Belgium', select = c("sex","year","prop_diabetes"))

# Print plot for each country
ggplot(data = BelgiumPlot, mapping = aes(x = year, y =prop_diabetes, color =sex)) + geom_line() + labs(title ="Evolution of diabetes prevalence in Belgium", x= "Year", y="Diabetes prevalence") + scale_y_continuous(labels = scales::percent)
```

A decrease over time for women but not for men for :

* Austria  
* Malta  
* Netherlands  
* Germany  
* Italy   
* Spain  
* Switzerland 

We take Austria as an example :
```{r}
# Taking subset of data table for the chosen country
AustriaPlot <- subset(Diabetes_EU2,country == 'Austria', select = c("sex","year","prop_diabetes"))

# Print plot for each country
ggplot(data = AustriaPlot, mapping = aes(x = year, y =prop_diabetes, color =sex)) + geom_line() + labs(title ="Evolution of diabetes prevalence in Austria", x= "Year", y="Diabetes prevalence") + scale_y_continuous(labels = scales::percent)
```

In other European countries, the prevalence of diabetes is increasing (at different paces) over time.
We take Croatia as an example :

```{r}
# Taking subset of data table for the chosen country
CroatiaPlot <- subset(Diabetes_EU2,country == 'Croatia', select = c("sex","year","prop_diabetes"))

# Print plot for each country
ggplot(data = CroatiaPlot, mapping = aes(x = year, y =prop_diabetes, color =sex)) + geom_line() + labs(title ="Evolution of diabetes prevalence in Croatia", x= "Year", y="Diabetes prevalence") + scale_y_continuous(labels = scales::percent)
```


## Plotting Diabetes against each type of macronutrient consumption

Finally, we want to plot the relation between the diabetes prevalence against the total calorie consumption as well as each type of macronutrient consumed.

```{r}
# Diabetes vs total calories
ggplot(data = GDP_diabetes_cal) +
  geom_point(mapping = aes(x = total_consumption, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = total_consumption, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = total_consumption, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = total_consumption, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs total calories per day", x="Total calories per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)

# Diabetes vs calories from animal protein
ggplot(data = GDP_diabetes_cal) +
  geom_point(mapping = aes(x = cal_prot_animal, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = cal_prot_animal, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = cal_prot_animal, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = cal_prot_animal, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs daily calories from animal protein", x="Total calories from animal protein per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)

# Diabetes vs calories from plant protein
ggplot(data = GDP_diabetes_cal) +
  geom_point(mapping = aes(x = cal_prot_plant, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = cal_prot_plant, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = cal_prot_plant, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = cal_prot_plant, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs daily calories from plant protein", x="Total calories from plant protein per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)

# Diabetes vs calories from carbohydrates
ggplot(data = GDP_diabetes_cal) +
  geom_point(mapping = aes(x = cal_carbs, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = cal_carbs, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = cal_carbs, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = cal_carbs, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs daily calories from carbohydrates", x="Total calories from carbohydrates per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)

# Diabetes vs calories from fat
ggplot(data = GDP_diabetes_cal) +
  geom_point(mapping = aes(x = cal_fat, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = cal_fat, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = cal_fat, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = cal_fat, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs daily calories from fat", x="Total calories from fat per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)
```

We can see a negative trend for the total consumption, the calories from animal protein and the calories from fat. We can observe a positive trend against calories from plant protein. For protein from carbohydrates, we can see a slighty positive trend for women.


## Plotting Diabetes against each type of macronutrient consumption (without outliers)

Now, since they affected our plots that included the GDP variable so much, we want to see if we have different trends when we remove our 5 outliers. 

```{r}
# Diabetes vs total calories
ggplot(data = GDP_diabetes_cal2) +
  geom_point(mapping = aes(x = total_consumption, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = total_consumption, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = total_consumption, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = total_consumption, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs total calories per day (without outliers)", x="Total calories per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)

# Diabetes vs calories from animal protein
ggplot(data = GDP_diabetes_cal2) +
  geom_point(mapping = aes(x = cal_prot_animal, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = cal_prot_animal, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = cal_prot_animal, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = cal_prot_animal, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs daily calories from animal protein (without outliers)", x="Total calories from animal protein per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)

# Diabetes vs calories from plant protein
ggplot(data = GDP_diabetes_cal2) +
  geom_point(mapping = aes(x = cal_prot_plant, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = cal_prot_plant, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = cal_prot_plant, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = cal_prot_plant, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs daily calories from plant protein (without outliers)", x="Total calories from plant protein per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)

# Diabetes vs calories from carbohydrates
ggplot(data = GDP_diabetes_cal2) +
  geom_point(mapping = aes(x = cal_carbs, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = cal_carbs, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = cal_carbs, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = cal_carbs, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs daily calories from carbohydrates (without outliers)", x="Total calories from carbohydrates per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)

# Diabetes vs calories from fat
ggplot(data = GDP_diabetes_cal2) +
  geom_point(mapping = aes(x = cal_fat, y = prop_men_diabetes, color = "blue")) +
  geom_smooth(mapping = aes(x = cal_fat, y = prop_men_diabetes, color = "blue"), formula = y ~ x, method = "lm") +
  geom_point(mapping = aes(x = cal_fat, y = prop_women_diabetes, color = "red")) +
  geom_smooth(mapping = aes(x = cal_fat, y = prop_women_diabetes, color = "red"), formula = y ~ x, method = "lm") +
  labs(title="Diabetes vs daily calories from fat (without outliers)", x="Total calories from fat per day", y="Diabetes prevalence") +
  scale_color_manual(labels = c("Men", "Women"), values = c("blue", "red")) +
  scale_y_continuous(labels = scales::percent)

```

Without our 5 outliers, we observe not much change in the trend of each type of calories consumed apart for carbohydrates where the trend changes for men and become slightly positive.
