---
title: "European countries, their dietary habits and diabetes prevalence"
output: html_notebook
---
## Introduction 

# Data
# 2 Data

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```


# Our research questions
Do European countries that have higher GDPs really have lower diabetes prevalence ? Do European countries that have higher GDPs consume less calories ? How do the proportions of macronutrients (animal protein/plant protein/fat/carbohydrates) consumed differ between richer and poorer governments ? And how do these differences relate to the diabetes prevalence in these countries ?  
 What is the typical diet that can be observed in richer states that relates to lower diabetes prevalence ?

# Analysis 

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/Code.R"))
```
1.Do European countries that have higher GDPs really have lower diabetes prevalence ?

```{r}
cor()
```


2.Do European countries that have higher GDPs consume less calories ?

3.How do the proportions of macronutrients (animal protein/plant protein/fat/carbohydrates) consumed differ between richer and poorer governments ?

4.And how do these differences relate to the diabetes prevalence in these countries ?

5.What is the typical diet that can be observed in richer states that relates to lower diabetes prevalence ?




Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.
```{r}
plot(cars)
```
When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.